---
title: "Project 1"
author: "Melody"
date: "2022-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r more setup, include=FALSE}
library(pacman)
p_load(ggplot2, tidyverse, here, tidytuesdayR, gapminder, stringr, magrittr)

chocolate <- readRDS(here("data","chocolate.RDS"))
as_tibble(chocolate)
chocolate <- as.data.frame(chocolate)

```

#### Part 1: Exploring Data

1.  Make a histogram of the rating scores to visualize overall distribution of scores, change the bins, pick the one you like best, and explain the difference among the bins and why you chose the one you did.

```{r histograms}
## default (30 bins)
ggplot(data=chocolate, aes(x=rating)) +
  geom_histogram(color="lightblue", fill="lightblue")

## 10 bins
ggplot(data=chocolate, aes(x=rating)) +
  geom_histogram(bins=10, color="salmon", fill="salmon")

## 15 bins
ggplot(data=chocolate, aes(x=rating)) +
  geom_histogram(bins=15, color="red", fill="red")

## 20 bins
ggplot(data=chocolate, aes(x=rating)) +
  geom_histogram(bins=20, color="lightpink", fill="lightpink")

## 25 bins
ggplot(data=chocolate, aes(x=rating)) +
  geom_histogram(bins=25, color="deeppink", fill="deeppink")

```

> I personally like the histogram with 15 bins (colored red) because it doesn't feel as disjointed as the histograms with more bins and so is visually nicer.

2.  How many reviews come from each country of bean origin?

```{r beans}
print(table(chocolate$country_of_bean_origin))
# is this asking how many?

```

3.  What is average rating scores from reviews of chocolate bars that have Ecuador as country_of_bean_origin in this dataset? For this same set of reviews, also calculate (1) the total number of reviews and (2) the standard deviation of the rating scores. Your answer should be a new data frame with these three summary statistics in three columns. Label the name of these columns mean, sd, and total.

```{r Ecuador}
ec <- chocolate %>%
  filter(country_of_bean_origin=="Ecuador")
nrow(ec)
mean(ec$rating)
sd(ec$rating, na.rm=TRUE)

## new dataframe with summary statistics
mean_rating <- c(mean(ec$rating))
number_reviews <- c(nrow(ec))
stdev_ratings <- c(sd(ec$rating))

df <- data.frame(mean_rating, number_reviews, stdev_ratings)
df

```

> There are 219 total reviews on chocolate made from beans from Ecuador. The average rating for beans from Ecuador is 3.16. The standard deviation of ratings is 0.512. Please see the new dataframe.

4.  Which country makes the best chocolate (or has the highest ratings on average) with beans from Ecuador?

```{r best beans}
a <- ec %>%
  group_by(company_location) %>%
  mutate(av_rating=mean(rating)) %>%
  distinct(av_rating) 
max(a$av_rating)
a1 <- a %>%
  filter(av_rating==3.8125)
a1

```

> The country that makes the best chocolate with beans from Ecuador with the highest average ratings is Australia.

5.  Calculate the average rating across all country of origins for beans. Which top 3 countries have the highest ratings on average?

```{r top three}
b <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  mutate(av_rating=mean(rating)) %>%
  distinct(av_rating) %>%
  arrange(desc(av_rating))
b

```

> The three countries of origin with the highest average ratings are Tobago, Sao Tome & Principe, and China in that order.

6.  Following up on the previous problem, now remove any countries of bean origins that have less than 10 chocolate bar reviews. Now, which top 3 countries have the highest ratings on average?

```{r best beans part two}
ch <- data.frame(table(chocolate$country_of_bean_origin))
ch <- ch %>%
  rename(country_of_bean_origin = Var1)

c <- left_join(chocolate, ch, by="country_of_bean_origin")  

c <- c %>%
  filter(!(Freq<=10)) %>%
  group_by(country_of_bean_origin) %>%
  mutate(av_rating=mean(rating)) %>%
  distinct(av_rating) %>%
  arrange(desc(av_rating))
c

```

> After removing countries of origin with fewer than 10 chocolate bar reviews, the three countries with the highest ratings on average are Congo, Cuba, and Vietnam, in that order.

7. Relationships between percent chocolate and ratings
```{r percent chocolate and ratings}
## 1. Countries of bean origin with at least 50 reviews
ch <- data.frame(table(chocolate$country_of_bean_origin))
ch <- ch %>%
  rename(country_of_bean_origin = Var1)

d <- left_join(chocolate, ch, by="country_of_bean_origin")
d <- d %>%
  filter(Freq>=50) 

## 2. Chocolate percentage
pc = paste(d$cocoa_percent)
as.numeric(sub("%", "", pc))

## 3. Refactoring
d <- d %>%
  mutate(percent=case_when(
    cocoa_percent<60 ~ "<60%",
    cocoa_percent>=60 & cocoa_percent<70 ~ "60%-70%",
    cocoa_percent>=70 & cocoa_percent<90 ~ "70%-90%",
    cocoa_percent>=90 ~ ">90%"
  ))

d$percent <- factor(x=d$percent, levels=c("<60%","60%-70%","70%-90%",">90%"), 
                    labels=c("<60%","60%-70%","70%-90%",">90%"))

## 4. Boxplots
ggplot(data=d, aes(x=percent, y=rating)) +
  geom_boxplot() +
  facet_wrap(~country_of_bean_origin)

```

> Of the four cocoa percentage categories, the two percentage categories that appear to be the most popular on average are 60%-70% and 70%-90%. This is not a consistent observation across the included countries, but most countries' most popular chocolate category is one of the two.


#### Part 2: Join two datasets together
```{r joining}
## prep
y <- data.frame(table(chocolate$country_of_bean_origin))
y <- y %>%
  rename(country_of_bean_origin = Var1)

e <- left_join(chocolate, y, by="country_of_bean_origin")
e <- e %>%
  filter(Freq>=10) %>%
  filter(!(country_of_bean_origin=="Blend")) %>%
  rename(country=country_of_bean_origin)
table(e$country_of_bean_origin)

z <- gapminder %>%
  pivot_longer(-c(country:year), names_to="stats", values_to="value") %>%
  mutate(country=recode(country, "United States"="U.S.A.")) %>%
  mutate(country=recode(country, "Congo, Dem. Rep."="Congo")) %>%
  select(country, continent)

## filling in the later NAs
## countries with NA continents: Belize (Americas), Congo (Africa), St. Lucia (Americas), Fiji (Oceania), Grenada (Americas), Papua New Guinea (Oceania), Sao Tome (Africa), Solomon Islands (Oceania), Trinidad (Americas), Vanuatu (Oceania)
country <- c("Belize", "St. Lucia","Fiji","Grenada","Papua New Guinea","Sao Tome","Solomon Islands","Trinidad","Vanuatu")
continent <- c("Americas", "Americas","Oceania","Americas","Oceania","Africa","Oceania",
               "Americas","Oceania")
df <- data.frame(country, continent)
v <- rbind(z, df)

table(v$country)

## joining and checking
x <- left_join(e, v, by="country")
table(x$country)
table(x$continent)

## violin plot
ggplot(data=x, aes(x=continent, y=rating)) +
  geom_violin(color="darkseagreen", fill="darkseagreen3") +
  geom_boxplot(width=0.1) +
  theme_minimal() +
  labs(title="Distribution of Chocolate Ratings per Continent", x="Rating (out of 4)",
       y="Continent", caption="Europe was not the source of any beans so was excluded")

```


#### Part 3: Wide to long data
```{r wide to long}
table(chocolate$ingredients)

## new set of columns part 1
f <- chocolate %>%
  mutate(beans=1) %>%
  mutate(sugar=ifelse(ingredients=="2- B,S" | ingredients=="3- B,S,C" | ingredients=="3- B,S,L" |
                        ingredients=="3- B,S,V" | ingredients=="4- B,S,C,L" | 
                        ingredients=="4- B,S,C,Sa" | ingredients=="4- B,S,C,V" |
                        ingredients=="4- B,S,V,L" | ingredients=="5- B,S,C,L,Sa" |
                        ingredients=="5- B,S,C,V,L" | ingredients=="5-B,S,C,V,Sa" |
                        ingredients=="6-B,S,C,V,L,Sa", 
                      1, 0)) %>%
  mutate(cocoa_butter=ifelse(ingredients=="2- B,C" | ingredients=="3- B,S*,C" | 
                               ingredients=="3- B,S,C" | ingredients=="4- B,S*,C,L" |
                               ingredients=="4- B,S*,C,Sa" | ingredients=="4- B,S*,C,V" |
                               ingredients=="4- B,S,C,L" | ingredients=="4- B,S,C,Sa" |
                               ingredients=="4- B,S,C,V" | ingredients=="4- B,S,V,L" |
                               ingredients=="5- B,S,C,L,Sa" | ingredients=="5- B,S,C,V,L" |
                               ingredients=="5-B,S,C,V,Sa" | ingredients=="6-B,S,C,V,L,Sa", 
                             1, 0)) %>%
  mutate(vanilla=ifelse(ingredients=="3- B,S,V" | ingredients=="4- B,S*,C,V" | 
                          ingredients=="4- B,S*,V,L" | ingredients=="4- B,S,C,V" | 
                          ingredients=="4- B,S,V,L" | ingredients=="5- B,S,C,V,L" |
                          ingredients=="5-B,S,C,V,Sa" | ingredients=="6-B,S,C,V,L,Sa",
                        1, 0)) %>%
  mutate(letchin=ifelse(ingredients=="3- B,S,L" | ingredients=="4- B,S*,C,L" |
                          ingredients=="4- B,S*,V,L" | ingredients=="4- B,S,C,L" |
                          ingredients=="4- B,S,V,L" | ingredients=="5- B,S,C,L,Sa" |
                          ingredients=="5- B,S,C,V,L" | ingredients=="6-B,S,C,V,L,Sa",
                        1, 0)) %>%
  mutate(salt=ifelse(ingredients=="3- B,S*,Sa" | ingredients=="4- B,S*,C,Sa" | 
                       ingredients=="4- B,S,C,Sa" | ingredients=="5- B,S,C,L,Sa" |
                       ingredients=="5-B,S,C,V,Sa" | ingredients=="6-B,S,C,V,L,Sa",
                     1, 0))

## new set of columns part 2
f$char_cocoa <- str_detect(f$most_memorable_characteristics, "cocoa")
f %<>% mutate(char_cocoa=ifelse(char_cocoa=="TRUE",1,0))
f$char_sweet <- str_detect(f$most_memorable_characteristics, "sweet")
f %<>% mutate(char_sweet=ifelse(char_sweet=="TRUE",1,0))
f$char_nutty <- str_detect(f$most_memorable_characteristics, "nutty")
f %<>% mutate(char_nutty=ifelse(char_nutty=="TRUE",1,0))
f$char_creamy <- str_detect(f$most_memorable_characteristics, "creamy")
f %<>% mutate(char_creamy=ifelse(char_creamy=="TRUE",1,0))
f$char_roasty <- str_detect(f$most_memorable_characteristics, "roasty")
f %<>% mutate(char_roasty=ifelse(char_roasty=="TRUE",1,0))
f$char_earthy <- str_detect(f$most_memorable_characteristics, "earthy")
f %<>% mutate(char_earthy=ifelse(char_earthy=="TRUE",1,0))

## mean value across all reviews per year
g <- f %>%
  select(review_date, beans, sugar, cocoa_butter, vanilla, letchin, salt, char_cocoa, char_sweet,
         char_nutty, char_creamy, char_roasty, char_earthy) %>%
  group_by(review_date) %>%
  mutate(beans_mean=mean(beans)) %>%
  mutate(sugar_mean=mean(sugar)) %>%
  mutate(cocoa_butter_mean=mean(cocoa_butter)) %>%
  mutate(vanilla_mean=mean(vanilla)) %>%
  mutate(letchin_mean=mean(letchin)) %>%
  mutate(salt_mean=mean(salt)) %>%
  mutate(char_cocoa_mean=mean(char_cocoa)) %>%
  mutate(char_sweet_mean=mean(char_sweet)) %>%
  mutate(char_nutty_mean=mean(char_nutty)) %>%
  mutate(char_creamy_mean=mean(char_creamy)) %>%
  mutate(char_roasty_mean=mean(char_roasty)) %>%
  mutate(char_earthy_mean=mean(char_earthy))
g <- g %>%
  select(review_date, beans_mean, sugar_mean, cocoa_butter_mean, vanilla_mean, letchin_mean,
         salt_mean, char_cocoa_mean, char_sweet_mean, char_nutty_mean, char_creamy_mean,
         char_roasty_mean, char_earthy_mean) %>%
  rename(beans=beans_mean,
         sugar=sugar_mean,
         cocoa_butter=cocoa_butter_mean,
         vanilla=vanilla_mean,
         letchin=letchin_mean,
         salt=salt_mean,
         char_cocoa=char_cocoa_mean,
         char_sweet=char_sweet_mean,
         char_nutty=char_nutty_mean,
         char_creamy=char_creamy_mean,
         char_roasty=char_roasty_mean,
         char_earthy=char_earthy_mean)

h <- g %>%
  pivot_longer(-review_date, names_to="feature", values_to="mean_scores") %>%
  distinct(review_date, feature, mean_scores)
print(h)

```

```{r alternative method of making new part 1 columns, include=FALSE}
fiwb <- chocolate %>%
  mutate(beans=case_when(grepl(pattern="B", x=ingredients)~"1",
                         TRUE~"0")) %>%
  mutate(sugar=case_when(grepl(pattern="S", x=ingredients)~"1",
                         TRUE~"0")) %>%
  mutate(cocoa_butter=case_when(grepl(pattern="C", x=ingredients)~"1",
                                TRUE~"0")) %>%
  mutate(vanilla=case_when(grepl(pattern="V", x=ingredients)~"1",
                           TRUE~"0")) %>%
  mutate(letchin=case_when(grepl(pattern="L", x=ingredients)~"1",
                           TRUE~"0")) %>%
  mutate(salt=case_when(grepl(pattern="Sa", x=ingredients)~"1",
                        TRUE~"0"))

```



#### Part 4: Data visualization
```{r data vis}
i <- ggplot(data=h, aes(review_date, mean_scores)) +
  geom_point() +
  geom_smooth(method="loess", se=FALSE, color="salmon") +
  facet_wrap(~feature) +
  theme_minimal() +
  labs(x="Year of review", y="Average score (out of 1)", 
       title="Change in average characteristic score over time", 
       subtitle="Average characteristic scores for chocolate generally stay constant with time",
       caption="Mel Xiao")
i

```


#### Part 5: Worst plot
```{r ugly plot}
j <- ggplot(data=chocolate, aes(x=review_date, y=ref)) +
  geom_line() +
  geom_path(lineend="butt",linejoin="round",linemitre=1, color="limegreen") +
  ylim(-5,80000) +
  labs(title="Title3", x="yeaR", y="amount", caption="please give me an A")
j

```

> 1. The y-axis is the reference number for each chocolate review, which is a wildly uninformative thing to plot over time (it does not change over time). 
> 2. Geom_path normally connects the pieces of data as they appear in the data. Here, the path is a disaster and we cannot see where it connects to or from anything.
> 3. The path is also lime green because in R, the color "limegreen" does not look great. The color is also not informative for anything (like grouping).
> 4. The y axis is scaled way too far out and we can't see any of the information that we actually have.
> 5. The title tells us nothing about the data we are looking at or what the graph is trying to depict.
> 6. The x and y axis labels also don't tell us about what units we are working with or even what the axes represent; for example, what does "amount" mean? And "year" is spelled weirdly.
> 7. The caption asks for you to give me an A, which does not relate to the assignment or the results and can probably be considered pandering. 
>
> When I am trying to make an actually good graph (respective to the above customizations):
> 1. I will look for a y-axis that actually changes over time, if my x-axis is time, so that the results actually make sense and I can track changes.
> 2. I won't add extra geoms that I don't need, and if I ever do need geom_path I will use it with data where it makes sense to track the order in which data appears.
> 3. I will use color more intelligently, such as to distinguish between groups; I will also use more visually appealing colors.
> 4. My y-axis can and will be scaled properly so that all the data is encompassed but is not so far out that the data is hard to see.
> 5. I will give my plot a coherent and useful title that tells me what the graph is actually of.
> 6. X and Y axis labels should tell me what the axis is representing and in what units.
> 7. The caption should be used for something useful, like my name or extra explanations.


#### Part 6: Better plot
```{r better plot}
k <- ggplot(data=chocolate, aes(x=review_date, y=rating, group=review_date)) +
  geom_boxplot() +
  labs(title="Distribution of chocolate scores across all countries and over time",
         x="Year of Review", y="Rating (out of 4)") +
  theme_minimal()
k

```

> 1. I removed the fill. Originally, having the fill was meant to show a gradient from earlier to later dates, but the fill was not actually informative; we can tell from the x-axis that we are looking at changes over time.
> 2. I changed the x-axis label to "Year of Review," which is a more informative and useful thing to name the x-axis.
> 3. I changed the y-axis label to "Score (out of 4)" so that we know what the ratings actually mean and what the scale of it is.
> 4. I added a title to the graph so we know what it is about.
> 5. I changed the geom to geom_boxplot because it is more informative than a violin plot in showing distribution data.
> 6. I left the year of review as a numeric variable because it is inherently a numeric piece of data that increases/has an ordering. 
> 7. I changed the theme to minimal so that the boxplots are more easily visible, especially the outliers.




